---
title: "mouse gastrulation"
output: html_notebook
---
 

```{r}
suppressPackageStartupMessages({
library(MouseGastrulationData)
library(SingleCellExperiment)
library(scater)
library(scran)
library(patchwork)
library(pheatmap)
library(EnsDb.Mmusculus.v79)  
library(viridis)
  library(magrittr)
  library(dplyr)
  library(ComplexHeatmap)
  library(circlize)
})
set.seed(100)
```


#load gastrulation data from E8.0 and 8.5

```{r}
AtlasSampleMetadata
sce <- EmbryoAtlasData(samples = c(33,34,35,36,37))

#remove stripped and doublet cells
inx <- which(!sce$doublet & !sce$stripped)
sce <- sce[,inx]

#add logcounts
sce <- logNormCounts(sce)

sce$NM <- ifelse(sce$celltype =="Spinal cord" | sce$celltype =="Rostral neurectoderm" | sce$celltype =="Forebrain/Midbrain/Hindbrain" ,"neuronal",
                 ifelse(sce$celltype =="Intermediate mesoderm" | sce$celltype =="Paraxial mesoderm" | sce$celltype =="Pharyngeal mesoderm"
                        | sce$celltype =="Mesenchyme" | sce$celltype =="Endothelium" | sce$celltype =="Somitic mesoderm" | sce$celltype =="Caudal Mesoderm" | sce$celltype =="Notochord" | sce$celltype =="Cardiomyocytes" | sce$celltype =="Mixed mesoderm" | sce$celltype =="Blood progenitors 1" | sce$celltype =="Blood progenitors 2" | sce$celltype =="Erythroid1" | sce$celltype =="Erythroid2" | sce$celltype =="Erythroid3" ,"mesodermal",
                        ifelse(sce$celltype =="NMP","NMP",
                               ifelse( sce$celltype =="Neural crest","Neural_crest",
                                       ifelse(sce$celltype=="Gut" | sce$celltype=="Def. endoderm" | sce$celltype=="Visceral endoderm" | sce$celltype=="Parietal endoderm","endodermal",
                                              ifelse(sce$celltype == "Surface ectoderm", "ectodermal","other"))))))
table(sce$NM)

```

# plot UMAP

```{r,fig.width=15,fig.height=8}
x <- data.frame(reducedDim(sce, "umap"))
colnames(x) <- c("dim1","dim2")

ggplot(x,aes(x=dim1,y=dim2,col=colData(sce)$celltype)) + geom_point() + 
               facet_wrap(vars(colData(sce)$stage)) + theme_classic() 
```

# load Bra-chip node genes

```{r}

node.genes <- readRDS("../fig6_with_Uniprot.rds")
View(node.genes)
#select the genes that are TFs, have a node info, and a mouse gene ID
node.genes2 <- node.genes[!is.na(node.genes$Lineage) & grepl("ENS",node.genes$Gene) & node.genes$TF=="TF",]
View(node.genes2)

#correct FoxA and Tbx2/3 genes according to trees
node.genes2$Lineage[c(grep("Foxa",node.genes2$gene_name),grep("foxa",node.genes2$gene_name))] <- c("Lineage2","Lineage2")
Foxa1 <- c("Foxa1","ENSMUSG00000035451","TF","Lineage2","NA","NA","NA")
node.genes2 <- rbind(node.genes2,Foxa1)

node.genes2$Lineage[grep("Tbx2",node.genes2$gene_name)] <- c("Lineage2","Lineage2")

#remove duplicated ones
length(unique(node.genes2$Gene))
dup.genes <- node.genes2$Gene[duplicated(node.genes2$Gene)]
View(node.genes2[node.genes2$Gene %in% dup.genes,])

remove.rows <- numeric(0)
for (i in seq_along(dup.genes)){
  remove.rows <- c(remove.rows,max(which(node.genes2$Gene==dup.genes[i])))
}
node.genes3 <- node.genes2[-remove.rows,]
```

# select node genes from sce object

```{r}
node1.inx <- which(rowData(sce)$ENSEMBL %in% node.genes3$Gene[node.genes3$Lineage=="Lineage1"])
rowData(sce)[node1.inx,]

node2.inx <- which(rowData(sce)$ENSEMBL %in% node.genes3$Gene[node.genes3$Lineage=="Lineage2"])
node3.inx <- which(rowData(sce)$ENSEMBL %in% node.genes3$Gene[node.genes3$Lineage=="Lineage3"])
node4.inx <- which(rowData(sce)$ENSEMBL %in% node.genes3$Gene[node.genes3$Lineage=="Lineage4"])
node5.inx <- which(rowData(sce)$ENSEMBL %in% node.genes3$Gene[node.genes3$Lineage=="Lineage5"])

```


# take the sum across logcounts per node and make violin plots across cells per node

```{r}
#average the logcounts acorss all genes in a node
node1.logcts <- colSums(logcounts(sce)[node1.inx,])
node2.logcts <- colSums(logcounts(sce)[node2.inx,])
node3.logcts <- colSums(logcounts(sce)[node3.inx,])
node4.logcts <- colSums(logcounts(sce)[node4.inx,])
node5.logcts <- colSums(logcounts(sce)[node5.inx,])


#combine in a df
nodes.df <- data.frame(celltype=sce$celltype,germ_layer=sce$NM,node1=node1.logcts,node2=node2.logcts,node3=node3.logcts,node4=node4.logcts,node5=node5.logcts)

p1 <- ggplot(nodes.df,aes(x=celltype,y=node1,fill=germ_layer)) + geom_violin() + theme_bw() + coord_flip() + scale_fill_manual(values=c("#f03b20","#feb24c","#984ea3","#e6f5d0","#7bccc4","#a6cee3","#fde0ef"))
p2 <- ggplot(nodes.df,aes(x=celltype,y=node2,fill=germ_layer)) + geom_violin() + theme_bw() + coord_flip() + scale_fill_manual(values=c("#f03b20","#feb24c","#984ea3","#e6f5d0","#7bccc4","#a6cee3","#fde0ef"))
p3 <- ggplot(nodes.df,aes(x=celltype,y=node3,fill=germ_layer)) + geom_violin() + theme_bw() + coord_flip() + scale_fill_manual(values=c("#f03b20","#feb24c","#984ea3","#e6f5d0","#7bccc4","#a6cee3","#fde0ef"))
p4 <- ggplot(nodes.df,aes(x=celltype,y=node4,fill=germ_layer)) + geom_violin() + theme_bw() + coord_flip() + scale_fill_manual(values=c("#f03b20","#feb24c","#984ea3","#e6f5d0","#7bccc4","#a6cee3","#fde0ef"))
p5 <- ggplot(nodes.df,aes(x=celltype,y=node5,fill=germ_layer)) + geom_violin() + theme_bw() + coord_flip() + scale_fill_manual(values=c("#f03b20","#feb24c","#984ea3","#e6f5d0","#7bccc4","#a6cee3","#fde0ef"))
#+ geom_boxplot(width=0.1, color="white", alpha=0.5)
p1 + p2 + p3 + p4 + p5
ggsave("../node_genes_sum_logcounts_per_celltype_violonplots_v5_FoxA_Tbx23_corrected.svg",device="svg",height=30,width=30)
```


# generate pseudo bulks per mesodermal, endodermal, and neuronal celltype

```{r}
summed <- aggregateAcrossCells(sce, 
    id=colData(sce)[,c("NM")])

#calculate normalization factor
norm.fact <- colSums(counts(summed))/1000000

counts.norm <- t(t(counts(summed))/norm.fact)
```

#make a ratio of all mesodermal vs neuronal celltypes for each gene and color them by this ratio, compare nodes

```{r}

inx.node.genes <- which(rowData(summed)$ENSEMBL %in% node.genes3$Gene)
summed.filtered <- data.frame(counts.norm[inx.node.genes,])


summed.filtered$logFC_N <- log2((summed.filtered$neuronal+1)/(summed.filtered$mesodermal+1))
summed.filtered$logFC_E <- log2((summed.filtered$endodermal+1)/(summed.filtered$mesodermal+1))

summed.filtered <- cbind(summed.filtered,rowData(sce)[inx.node.genes,])
summed.filtered <- left_join(summed.filtered,node.genes3,by=c("ENSEMBL"="Gene"))

ggplot(summed.filtered,aes(x=Lineage,y=logFC_N)) + geom_boxplot(notch=FALSE,fill="#7bccc4") +theme_bw() + ylab("log2 neuronal/mesodermal gene expression [cpm]") + xlab("") +  scale_x_discrete(labels=c("Lineage1" = "node 1", "Lineage2" = "node 2",
                              "Lineage3" = "node 3","Lineage4"="node 4","Lineage5"="node 5")) + theme(axis.text.x = element_text(  
                           size=14, angle=45,vjust = 0.5),axis.text.y = element_text( size=14), axis.title.y = element_text(face="bold",size=14)) 

wilcox.test(summed.filtered$logFC_N[summed.filtered$Lineage=="Lineage2"],summed.filtered$logFC_N[summed.filtered$Lineage=="Lineage5"])
wilcox.test(summed.filtered$logFC_N[summed.filtered$Lineage=="Lineage2"],summed.filtered$logFC_N[summed.filtered$Lineage=="Lineage4"])
ggsave("../boxplot_neuronal_mesodermal_FC_Lineages1-5_not_cpm_filtered_v2_FoxA_Tbx23_corrected.svg",device = "svg",height=8,width=6)

ggplot(summed.filtered,aes(x=Lineage,y=logFC_E)) + geom_boxplot(notch=FALSE,fill="#feb24c") +theme_bw() + ylab("log2 neuronal/mesodermal gene expression [cpm]") + xlab("") +  scale_x_discrete(labels=c("Lineage1" = "node 1", "Lineage2" = "node 2",
                              "Lineage3" = "node 3","Lineage4"="node 4","Lineage5"="node 5")) + theme(axis.text.x = element_text(  
                           size=14, angle=45,vjust = 0.5),axis.text.y = element_text( size=14), axis.title.y = element_text(face="bold",size=14)) 
wilcox.test(summed.filtered$logFC_E[summed.filtered$Lineage=="Lineage2"],summed.filtered$logFC_E[summed.filtered$Lineage=="Lineage5"])
wilcox.test(summed.filtered$logFC_E[summed.filtered$Lineage=="Lineage2"],summed.filtered$logFC_E[summed.filtered$Lineage=="Lineage4"])
ggsave("../boxplot_endodermal_mesodermal_FC_Lineages1-5_not_cpm_filtered_v2_FoxA_Tbx23_corrected.svg",device = "svg",height=8,width=6)


#remove the genes with < 1 cpm in meso and neuro combined
summed.filtered2 <- summed.filtered[rowSums(summed.filtered[,c(3,5)]) > 1,]

wilcox.test(summed.filtered2$logFC[summed.filtered2$Lineage=="Lineage2"],summed.filtered2$logFC[summed.filtered2$Lineage=="Lineage5"])
wilcox.test(summed.filtered2$logFC[summed.filtered2$Lineage=="Lineage2"],summed.filtered2$logFC[summed.filtered2$Lineage=="Lineage4"])

ggplot(summed.filtered2,aes(x=Lineage,y=logFC)) + geom_boxplot(notch=FALSE,fill="#7bccc4") +theme_bw()


#remove the genes with < 1 cpm in meso and endo combined
summed.filtered2 <- summed.filtered[rowSums(summed.filtered[,c(3,2)]) > 1,]

wilcox.test(summed.filtered2$logFC[summed.filtered2$Lineage=="Lineage2"],summed.filtered2$logFC[summed.filtered2$Lineage=="Lineage5"])
wilcox.test(summed.filtered2$logFC[summed.filtered2$Lineage=="Lineage2"],summed.filtered2$logFC[summed.filtered2$Lineage=="Lineage4"])

ggplot(summed.filtered2,aes(x=Lineage,y=logFC)) + geom_boxplot(notch=FALSE,fill="#7bccc4") +theme_bw()


```

# cpm heatmap without filtering

```{r}
row.names(summed.filtered) <- summed.filtered$SYMBOL

library(circlize)
col_fun = colorRamp2(c(0, max(log2(summed.filtered[,1:4]+1))), c("white", "darkblue"))
svg("../heatmap_neuronal_mesodermal_cpms_Lineages1-5_Foxa_Tbx23_corrected.svg",height=18,width=8)
Heatmap(log2(summed.filtered[,1:6]+1),cluster_columns = FALSE,split = summed.filtered$Lineage,col=col_fun,name="log2 cpm",cluster_row_slices = FALSE)
dev.off()

```

#logFC heatmap without filtering, different color for neuronal and endodermal

```{r}
row.names(summed.filtered) <- summed.filtered$SYMBOL


col_fun = colorRamp2(c(-2,0, 2), c("#984ea3","white", "#7bccc4"))
summed.filteredlogFC <- matrix(summed.filtered[,8])
row.names(summed.filteredlogFC) <- row.names(summed.filtered)
H1 <- Heatmap(summed.filteredlogFC,cluster_columns = FALSE,split = summed.filtered$Lineage,col=col_fun,name="logFC neuronal/mesodermal",show_row_names = TRUE,cluster_row_slices = FALSE,show_row_dend = FALSE)

col_fun = colorRamp2(c(-2,0, 2), c("#984ea3","white", "#feb24c"))
summed.filteredlogFC <- matrix(summed.filtered[,9])
row.names(summed.filteredlogFC) <- row.names(summed.filtered)
H2 <- Heatmap(summed.filteredlogFC,cluster_columns = FALSE,split = summed.filtered$Lineage,col=col_fun,name="logFC endodermal/mesodermal",show_row_names = TRUE,cluster_row_slices = FALSE,show_row_dend = FALSE)

svg("../heatmap_neuronal_mesodermal_FC_Lineages1-5_not_cpm_filtered_Foxa_Tbx23_corrected_with_endoderm_2colors.svg",height=18,width=8)
H1 + H2
dev.off()
```

#logFC heatmap without filtering, different color for neuronal and endodermal, clustered by combined endo and neuronal

```{r}
row.names(summed.filtered) <- summed.filtered$SYMBOL

library(cluster)
summed.filteredlogFCall <- summed.filtered[,8:9]
dist.eu <- dist(summed.filteredlogFCall)
clustering.all <- hclust(dist.eu)
row.ordering <- clustering.all[[3]]

col_fun = colorRamp2(c(-2,0, 2), c("#984ea3","white", "#7bccc4"))
summed.filteredlogFC <- matrix(summed.filtered[,8])
row.names(summed.filteredlogFC) <- row.names(summed.filtered)
summed.filteredlogFC <- summed.filteredlogFC[row.ordering,]
H1 <- Heatmap(summed.filteredlogFC,cluster_columns = FALSE,cluster_rows= FALSE, split = summed.filtered$Lineage[row.ordering],col=col_fun,name="logFC neuronal/mesodermal",show_row_names = TRUE,cluster_row_slices = FALSE,show_row_dend = FALSE)

col_fun = colorRamp2(c(-2,0, 2), c("#984ea3","white", "#feb24c"))
summed.filteredlogFC <- matrix(summed.filtered[,9])
row.names(summed.filteredlogFC) <- row.names(summed.filtered)
summed.filteredlogFC <- summed.filteredlogFC[row.ordering,]
H2 <- Heatmap(summed.filteredlogFC,cluster_columns = FALSE,cluster_rows= FALSE, split = summed.filtered$Lineage[row.ordering],col=col_fun,name="logFC endodermal/mesodermal",show_row_names = TRUE,cluster_row_slices = FALSE,show_row_dend = FALSE)

svg("../heatmap_neuronal_mesodermal_FC_Lineages1-5_not_cpm_filtered_Foxa_Tbx23_corrected_with_endoderm_2colors_euclidian_clustering.svg",height=18,width=8)
H1 + H2
dev.off()
```

#logFC heatmap without filtering, different color for neuronal and endodermal, clustered seperatly

```{r}
row.names(summed.filtered) <- summed.filtered$SYMBOL

col_fun = colorRamp2(c(-2,0, 2), c("#984ea3","white", "#7bccc4"))
summed.filteredlogFC <- matrix(summed.filtered[,8])
row.names(summed.filteredlogFC) <- row.names(summed.filtered)
H1 <- Heatmap(summed.filteredlogFC,cluster_columns = FALSE,cluster_rows= TRUE, split = summed.filtered$Lineage,col=col_fun,name="logFC neuronal/mesodermal",show_row_names = FALSE,
              cluster_row_slices = FALSE,show_row_dend = FALSE,
              cell_fun = function(j, i, x, y, width, height, fill) {
            gp = gpar(col = "lightgrey", fill = NA))
        grid.text( row.names(summed.filteredlogFC)[i], x=x, y=y, gp = gpar(fontsize = 10))
        
})

summed.filtered$Uniprot_colors <- ifelse(summed.filtered$Uniprot_function =="Mesodermal","black","red")
H1 <- Heatmap(summed.filteredlogFC,cluster_columns = FALSE,cluster_rows= TRUE, split = summed.filtered$Lineage,col=col_fun,name="logFC neuronal/mesodermal",show_row_names = FALSE,
              cluster_row_slices = FALSE,show_row_dend = FALSE,
              cell_fun = function(j, i, x, y, width, height, fill) {
                grid.rect(x = x, y = y, width = width, height = height, 
            gp = gpar(col = summed.filtered$Uniprot_colors, fill = NA))
        grid.text( row.names(summed.filteredlogFC)[i], x=x, y=y, gp = gpar(fontsize = 10))
      

})



col_fun = colorRamp2(c(-2,0, 2), c("#984ea3","white", "#feb24c"))
summed.filteredlogFC <- matrix(summed.filtered[,9])
row.names(summed.filteredlogFC) <- row.names(summed.filtered)
H2 <- Heatmap(summed.filteredlogFC,cluster_columns = FALSE,cluster_rows= TRUE, split = summed.filtered$Lineage,col=col_fun,name="logFC endodermal/mesodermal",show_row_names = FALSE,
              cluster_row_slices = FALSE,show_row_dend = FALSE,
              cell_fun = function(j, i, x, y, width, height, fill) {
                grid.rect(x = x, y = y, width = width, height = height, 
            gp = gpar(col = "lightgrey", fill = NA))
        grid.text( row.names(summed.filteredlogFC)[i], x=x, y=y, gp = gpar(fontsize = 10))
})

svg("../heatmap_neuronal_mesodermal_FC_Lineages1-5_not_cpm_filtered_Foxa_Tbx23_corrected_2colors_v2.svg",height=19,width=3.7)
H1
dev.off()

svg("../heatmap_endodermal_mesodermal_FC_Lineages1-5_not_cpm_filtered_Foxa_Tbx23_corrected_2colors_v2.svg",height=19,width=3.7)
H2
dev.off()

```

#logFC heatmap without filtering, different color for neuronal and endodermal, clustered seperatly, with gene symbol and uniprot function insde 

```{r}
row.names(summed.filtered) <- paste(summed.filtered$SYMBOL,summed.filtered$Uniprot_function)
row.names(summed.filtered) <- gsub("NA","",row.names(summed.filtered))

col_fun = colorRamp2(c(-2,0, 2), c("#984ea3","white", "#7bccc4"))
summed.filteredlogFC <- matrix(summed.filtered[,8])
row.names(summed.filteredlogFC) <- row.names(summed.filtered)
H1 <- Heatmap(summed.filteredlogFC,cluster_columns = FALSE,cluster_rows= TRUE, split = summed.filtered$Lineage,col=col_fun,name="logFC neuronal/mesodermal",show_row_names = FALSE,
              cluster_row_slices = FALSE,show_row_dend = FALSE,
              cell_fun = function(j, i, x, y, width, height, fill) {
                
                  grid.rect(x = x, y = y, width = width, height = height, 
            gp = gpar(col = "lightgrey", fill = NA))  
               
        grid.text( row.names(summed.filteredlogFC)[i], x=x, y=y, gp = gpar(fontsize = 10))
        
})



col_fun = colorRamp2(c(-2,0, 2), c("#984ea3","white", "#feb24c"))
summed.filteredlogFC <- matrix(summed.filtered[,9])
row.names(summed.filteredlogFC) <- row.names(summed.filtered)
H2 <- Heatmap(summed.filteredlogFC,cluster_columns = FALSE,cluster_rows= TRUE, split = summed.filtered$Lineage,col=col_fun,name="logFC endodermal/mesodermal",show_row_names = FALSE,
              cluster_row_slices = FALSE,show_row_dend = FALSE,
              cell_fun = function(j, i, x, y, width, height, fill) {
                grid.rect(x = x, y = y, width = width, height = height, 
            gp = gpar(col = "lightgrey", fill = NA))
        grid.text( row.names(summed.filteredlogFC)[i], x=x, y=y, gp = gpar(fontsize = 10))
})

svg("../heatmap_neuronal_mesodermal_FC_Lineages1-5_not_cpm_filtered_Foxa_Tbx23_corrected_2colors_v2.svg",height=19,width=4)
H1
dev.off()

svg("../heatmap_endodermal_mesodermal_FC_Lineages1-5_not_cpm_filtered_Foxa_Tbx23_corrected_2colors_v2.svg",height=19,width=4)
H2
dev.off()

```

#logFC heatmap without filtering 

```{r}
row.names(summed.filtered) <- summed.filtered$SYMBOL
col_fun = colorRamp2(c(-2,0, 2), c("#d8b365","white", "#5ab4ac"))
#summed.filteredlogFC <- matrix(summed.filtered[,8])
summed.filteredlogFC <- summed.filtered[,8:9]

row.names(summed.filteredlogFC) <- row.names(summed.filtered)
svg("../heatmap_neuronal_mesodermal_FC_Lineages1-5_not_cpm_filtered_Foxa_Tbx23_corrected_with_endoderm.svg",height=18,width=6)
Heatmap(summed.filteredlogFC,cluster_columns = FALSE,split = summed.filtered$Lineage,col=col_fun,name="logFC",show_row_names = TRUE,cluster_row_slices = FALSE,show_row_dend = FALSE)
dev.off()
```

#logFC heatmap with filtering 

```{r}
row.names(summed.filtered2) <- summed.filtered2$SYMBOL
col_fun = colorRamp2(c(-2,0, 2), c("#d8b365","white", "#5ab4ac"))
summed.filtered2logFC <- matrix(summed.filtered2[,8])
row.names(summed.filtered2logFC) <- row.names(summed.filtered2)
svg("../heatmap_neuronal_mesodermal_FC_Lineages1-5.svg",height=18,width=6)
Heatmap(summed.filtered2logFC,cluster_columns = FALSE,split = summed.filtered2$Lineage,col=col_fun,name="logFC neuronal/mesodermal",show_row_names = TRUE,cluster_row_slices = FALSE,show_row_dend = FALSE)
dev.off()
```





# generate pseudo bulks per mesoerm and neuronal celltype, split by embryo stage

```{r}
summed <- aggregateAcrossCells(sce, 
    id=colData(sce)[,c("NM","stage")])

#calculate normalization factor
norm.fact <- colSums(counts(summed))/1000000

counts.norm <- t(t(counts(summed))/norm.fact)
```

#make a ratio of all mesodermal vs neuronal celltypes for each gene and color them by this ratio, compare nodes

```{r}

inx.node.genes <- which(rowData(summed)$ENSEMBL %in% node.genes3$Gene)
summed.filtered <- data.frame(counts.norm[inx.node.genes,])
colnames(summed.filtered) <- paste(summed$NM,summed$stage,sep="_")

summed.filtered$logFC_E8.0 <- log2((summed.filtered$neuronal_E8.0 + 1)/(summed.filtered$mesodermal_E8.0 + 1))
summed.filtered$logFC_E8.5 <- log2((summed.filtered$neuronal_E8.5 + 1)/(summed.filtered$mesodermal_E8.5 + 1))

summed.filtered <- cbind(summed.filtered,rowData(sce)[inx.node.genes,])
summed.filtered <- left_join(summed.filtered,node.genes3,by=c("ENSEMBL"="Gene"))

ggplot(summed.filtered,aes(x=Lineage,y=logFC_E8.0)) + geom_boxplot(notch=FALSE,fill="#7bccc4") +theme_bw()
wilcox.test(summed.filtered$logFC_E8.0[summed.filtered$Lineage=="Lineage2"],summed.filtered$logFC_E8.0[summed.filtered$Lineage=="Lineage5"])

ggplot(summed.filtered,aes(x=Lineage,y=logFC_E8.5)) + geom_boxplot(notch=FALSE,fill="#7bccc4") +theme_bw()
wilcox.test(summed.filtered$logFC_E8.5[summed.filtered$Lineage=="Lineage2"],summed.filtered$logFC_E8.5[summed.filtered$Lineage=="Lineage5"])

```



# cpm heatmap without filtering

```{r}
row.names(summed.filtered) <- summed.filtered$SYMBOL

library(circlize)
col_fun = colorRamp2(c(0, max(log2(summed.filtered[,1:4]+1))), c("white", "darkblue"))
svg("../heatmap_neuronal_mesodermal_cpms_Lineages1-5_stage_split.svg",height=18,width=14)
Heatmap(log2(summed.filtered[,1:13]+1),cluster_columns = FALSE,split = summed.filtered$Lineage,col=col_fun,name="log2 cpm",cluster_row_slices = FALSE)
dev.off()

```

#logFC heatmap without filtering 

```{r}
row.names(summed.filtered) <- summed.filtered$SYMBOL
col_fun = colorRamp2(c(-2,0, 2), c("#d8b365","white", "#5ab4ac"))
summed.filteredlogFC <- as.matrix(summed.filtered[,14:15])

row.names(summed.filteredlogFC) <- row.names(summed.filtered)
svg("../heatmap_neuronal_mesodermal_FC_Lineages1-5_not_cpm_filtered_stage_split.svg",height=18,width=8)
Heatmap(summed.filteredlogFC,cluster_columns = FALSE,split = summed.filtered$Lineage,col=col_fun,name="logFC neuronal/mesodermal",show_row_names = TRUE,cluster_row_slices = FALSE,show_row_dend = FALSE)
dev.off()
```


# generate pseudo bulks per mesoerm and neuronal celltype, split by sample

```{r}
summed <- aggregateAcrossCells(sce, 
    id=colData(sce)[,c("NM","sample")])

#calculate normalization factor
norm.fact <- colSums(counts(summed))/1000000

counts.norm <- t(t(counts(summed))/norm.fact)
```

#make a ratio of all mesodermal vs neuronal celltypes for each gene and color them by this ratio, compare nodes

```{r}

inx.node.genes <- which(rowData(summed)$ENSEMBL %in% node.genes3$Gene)
summed.filtered <- data.frame(counts.norm[inx.node.genes,])
colnames(summed.filtered) <- paste(summed$NM,summed$sample,sep="_")

summed.filtered$logFC_33 <- log2((summed.filtered$neuronal_33 + 1)/(summed.filtered$mesodermal_33 + 1))
summed.filtered$logFC_34 <- log2((summed.filtered$neuronal_34 + 1)/(summed.filtered$mesodermal_34 + 1))
summed.filtered$logFC_35 <- log2((summed.filtered$neuronal_35 + 1)/(summed.filtered$mesodermal_35 + 1))
summed.filtered$logFC_36 <- log2((summed.filtered$neuronal_36 + 1)/(summed.filtered$mesodermal_36 + 1))
summed.filtered$logFC_37 <- log2((summed.filtered$neuronal_37 + 1)/(summed.filtered$mesodermal_37 + 1))

summed.filtered <- cbind(summed.filtered,rowData(sce)[inx.node.genes,])
summed.filtered <- left_join(summed.filtered,node.genes3,by=c("ENSEMBL"="Gene"))

ggplot(summed.filtered,aes(x=Lineage,y=logFC_E8.0)) + geom_boxplot(notch=FALSE,fill="#7bccc4") +theme_bw()
wilcox.test(summed.filtered$logFC_E8.0[summed.filtered$Lineage=="Lineage2"],summed.filtered$logFC_E8.0[summed.filtered$Lineage=="Lineage5"])

ggplot(summed.filtered,aes(x=Lineage,y=logFC_E8.5)) + geom_boxplot(notch=FALSE,fill="#7bccc4") +theme_bw()
wilcox.test(summed.filtered$logFC_E8.5[summed.filtered$Lineage=="Lineage2"],summed.filtered$logFC_E8.5[summed.filtered$Lineage=="Lineage5"])

```



# cpm heatmap without filtering

```{r}
row.names(summed.filtered) <- summed.filtered$SYMBOL

library(circlize)
col_fun = colorRamp2(c(0, max(log2(summed.filtered[,1:4]+1))), c("white", "darkblue"))
svg("../heatmap_neuronal_mesodermal_cpms_Lineages1-5_sample_split.svg",height=18,width=25)
Heatmap(log2(summed.filtered[,1:31]+1),cluster_columns = FALSE,split = summed.filtered$Lineage,col=col_fun,name="log2 cpm",cluster_row_slices = FALSE)
dev.off()

```

#logFC heatmap without filtering 

```{r}
row.names(summed.filtered) <- summed.filtered$SYMBOL
col_fun = colorRamp2(c(-2,0, 2), c("#d8b365","white", "#5ab4ac"))
summed.filteredlogFC <- as.matrix(summed.filtered[,32:36])

row.names(summed.filteredlogFC) <- row.names(summed.filtered)
svg("../heatmap_neuronal_mesodermal_FC_Lineages1-5_not_cpm_filtered_sample_split.svg",height=18,width=12)
Heatmap(summed.filteredlogFC,cluster_columns = FALSE,split = summed.filtered$Lineage,col=col_fun,name="logFC neuronal/mesodermal",show_row_names = TRUE,cluster_row_slices = FALSE,show_row_dend = FALSE)
dev.off()
```

